---
title: "Lecture 11 - Key"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

#### Hierarchical Regression
Recall the hierarchical normal model we used previously.
\begin{eqnarray*}
 p(y| \theta_j, \sigma^2) &=& normal(\theta_j,\sigma^2) \text{  within-group model}\\
 p(\theta_j|\mu,\tau) &=& normal(\mu,\tau^2) \text{  between-group model}
\end{eqnarray*}
This model allowed a different mean for each group (school), denoted $\theta_j$.
\vfill
Now returning to the motivating example, test scores within a school. Suppose there are other factors that affect test scores at the school, specifically how about the socio-economic standing of families in that school district.
\vfill
We can now write the model as:
\begin{eqnarray*}
 p(y|\tilde{x}_j, \theta_j, \sigma^2) &=& normal(\tilde{x}_j^T \tilde{\theta}_j,\sigma^2) \text{  within-group model}\\
 p(\tilde{\theta}_j|\tilde{\mu},\Sigma) &=& MVN(\tilde{\mu},\Sigma) \text{  between-group model}
\end{eqnarray*}
where $\tilde{x}_J$ is a vector of socio-economic information about school $j$, this could also include $1$ to account for an intercept.
\vfill
What priors to we need to fit this model?
\vfill
\begin{eqnarray*}
\sigma^2 & \sim & InvGamma(\nu_0/2,\nu_0\sigma_0^2/2)\\
\Sigma & \sim & InvWishart(\eta_0/2,\eta_0\tau_0^2/2)\\
\tilde{\mu} &\sim & MVN(\tilde{\mu}_0,\Lambda_0)
\end{eqnarray*}
\vfill
Similar to the hierarchical means model, we can obtain full conditional distributions for $\sigma^2$, $\Sigma_0$, $\tilde{\theta}$, and $\tilde{\mu}$. This allows us to use a Gibbs sampler to draw samples from the joint posterior distribution.
\newpage

### Exercise

1. Write out the model for a hierarchical regression setting. To keep it simple assume we are fitting a different intercept and slope (associated with square footage) for the different zip codes.
\vfill

\begin{eqnarray*}
y_{ij} &\sim& N(X_{ij}\widetilde{\theta_{i}}, \sigma^2)\\
\widetilde{\theta_{i}} &\sim& N(\mu_0, \Sigma_0),\\
\end{eqnarray*}

where $y_{ij}$ is the price of the $j^{th}$ house in the $i^{th}$ zip code, $\widetilde{\theta_{i}}$ is the vector of regression coefficients (slope, intercept) for zip code $i$.
\vfill
The following priors are specified for this setting.
\begin{eqnarray*}
\sigma^2 & \sim & InvGamma(\nu_0/2,\nu_0\sigma_0^2/2)\\
\Sigma & \sim & InvWishart(\eta_0/2,\eta_0\tau_0^2/2)\\
\tilde{\mu} &\sim & MVN(\tilde{\mu}_0,\Lambda_0)
\end{eqnarray*}

\vfill
##### Extract Data

```{r, fig.height=10, fig.width=8}
library(readr)
library(dplyr)
library(tidyr)
seattle <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat532/data/SeattleHousing.csv')
set.seed(11122018)

num.zips <- 10
num.houses <- 20
keep.zips <- sample(unique(seattle$zipcode), num.zips)

seattle.filter <- seattle %>% filter(zipcode %in% keep.zips) %>% group_by(zipcode) %>% 
  sample_n(num.houses) %>% arrange(zipcode) %>% select(price, sqft_living, zipcode, id) %>% 
  ungroup() %>% mutate(zipcode = as.factor(zipcode), house.num = rep(1:num.houses, num.zips))

price.wide <- seattle.filter %>% select(zipcode, price,house.num) %>% 
  spread(key = zipcode, value = price) %>% select(-house.num)

size.wide <- seattle.filter %>% select(zipcode, sqft_living,house.num) %>% 
  spread(key = zipcode, value = sqft_living) %>% select(-house.num)


library(ggplot2)
ggplot(data = seattle.filter, aes(y = price, x = sqft_living)) + geom_point() + 
  geom_smooth() + facet_wrap(~zipcode, nrow = 5, ncol = 2)
```
\vfill

\newpage

2. Compare and contrast the following two models

\vfill
```{r}
summary(lm(price~ sqft_living , data = seattle.filter))
```
\vfill
```{r}
library(rjags)
modelstring <- "
model {
    # Model
    for (zip in 1:num.zips) {
        for (house in 1:num.houses) {
            mu[house, zip] <- alpha + beta * (x[house,zip]);
            price.wide[house,zip] ~ dnorm(mu[house,zip], tau.price)
        }
    }
    # Priors
    alpha   ~ dnorm(0, 1/1e16);
    beta    ~ dnorm(0, 1/1e16);
    tau.price     ~ dgamma(.0001, .0001);
    
    # Transformations
    sigma.price    <- 1.0/sqrt(tau.price);
}
"
writeLines(modelstring, "model.txt")

Data <- list(
    num.zips =   num.zips,
    num.houses =    num.houses,
    price.wide =   price.wide,
    x = size.wide)
mod1 <- jags.model("model.txt", data=Data, n.chains=4, n.adapt=1000)

codaSamples = coda.samples( mod1 , variable.names=c("alpha","beta", 'sigma.price') ,
                            n.iter=10000)

summary(codaSamples)
```
\vfill
\newpage
3. Now again, compare and contrast the following two models.
\vfill
```{r}
summary(lm(price~ sqft_living + zipcode - 1, data = seattle.filter))
```

\vfill
```{r}
library(rjags)
modelstring <- "
model {
    # Model
    for (zip in 1:num.zips) {
        for (house in 1:num.houses) {
            mu[house, zip] <- alpha[zip] + beta * (x[house,zip]);
            price.wide[house,zip] ~ dnorm(mu[house,zip], tau.price)
        }
        alpha[zip] ~ dnorm(alpha.mu, alpha.tau);
    }
    # Priors
    alpha.mu   ~ dnorm(0, 1/1e16);
    beta    ~ dnorm(0, 1/1e16);
    tau.price     ~ dgamma(.0001, .0001);
    alpha.tau ~ dgamma(.00001, .00001);
    # Transformations
    alpha.sigma  <- 1.0/sqrt(alpha.tau);
    sigma.price    <- 1.0/sqrt(tau.price);
}
"
writeLines(modelstring, "model.txt")

Data <- list(
    num.zips =   num.zips,
    num.houses =    num.houses,
    price.wide =   price.wide,
    x = size.wide)
mod1 <- jags.model("model.txt", data=Data, n.chains=4, n.adapt=1000)

codaSamples = coda.samples( mod1 , variable.names=c("alpha","beta", 'sigma.price', 'alpha.mu') ,
                            n.iter=10000)

summary(codaSamples)
```

\vfill